---
title: "Venn diagram to show overlap between MCF7 ARCHS4, BRCA, and MCF7 GSE"
author: "Vy Tran"
date: "July 6, 2020"
output: html_document
---

```{r setup, include=FALSE}
filepath = "D:/vytran/OneDrive - JHSPH/Year 3/Evaluate-the-validity-of-the-breast-cancer-cell-line-MCF7-as-a-model-for-breast-cancer-research/"
setwd(filepath)
getwd()
options(stringsAsFactors=FALSE) 
```

```{r library}
library(WGCNA)
library(GEOquery)
library(stringr)
library(tidyverse)
library(VennDiagram)
```
batch_corrected_RNA_MCF7_voom: MCF7 data was download from ARCHS4, normalized using voom. Now we fitler for the top 10,000 genes to reduce noise. 
```{r}

load(file = "D:/vytran/OneDrive - JHSPH/Year 3/Evaluate-the-validity-of-the-breast-cancer-cell-line-MCF7-as-a-model-for-breast-cancer-research/batch_corrected_RNA_MCF7_ARCHS4.Rdata")

MCF7_ARCHS4 = t(batch_corrected_RNA_MCF7_voom[order(apply(batch_corrected_RNA_MCF7_voom,1,mad), decreasing = T)[1:10000],])
dim(MCF7_ARCHS4)
```
WGCNA_matrix_BRCA:  BRCA data was downloaded from TCGA as level 3 normalized data. The top 10,000 genes were filtered. 
```{r message=FALSE, warning=FALSE, paged.print=FALSE}
dim(WGCNA_matrix_BRCA)
```
Next we obtain the top 10000 genes in MCF7 GEO. The resulting data set is called MCF7_estrogen.

View the gene names of the three data sets:
```{r}
colnames(MCF7_ARCHS4)[1:10]
colnames(WGCNA_matrix_BRCA)[1:10]
colnames(MCF7_estrogen)[1:10]
```

Extract gene names for each data set: 
```{r}
gene_ARCHS4 = colnames(MCF7_ARCHS4)
gene_ARCHS4_1 = str_extract(gene_ARCHS4, "[^-]+")
```
View the result:
```{r}
gene_ARCHS4[1:10]
gene_ARCHS4_1[1:10]
```

Similarly, extract gene names for BRCA:
```{r}
gene_BRCA = str_extract(colnames(WGCNA_matrix_BRCA), "[^|]+")
head(gene_BRCA, 10)

```
Extract gene names for GEO:
```{r}
#\\b is a zero-length token that matches a word-boundary. This includes any non-word characters:
gene_GEO = str_extract(colnames(MCF7_estrogen), '\\b\\w+$')
head(gene_GEO, 10)

```

(NOTE: This chunk was not run.) For the GEO MCF7 data set, we convert Affymetrix probe to gene symbols using the limma package.
```{r}
#Load the limma package:
     library(limma)
     
    #Installing annotation packages. The appropriate packages for the specific Affymetrix platform can be found on Bioconductor: 
     #BiocManager::install(c("hgu133plus2cdf", "hgu133plus2.db"))
     library(hgu133plus2cdf)
     library(hgu133plus2.db)
     
    #Get gene symbols. Multiple entries will be collapsed and separated by semicolons, e.g. if geneID 001 corresponds to gene names 
    #"Name1" and "Name2", these will be collapsed into "Name1; Name2":
     probes = colnames(MCF7_estrogen)
     gene_GEO <- as.character(unlist(lapply(mget(probes,env=hgu133plus2SYMBOL),
     function (symbol) { return(paste(symbol,collapse="; ")) } )))
    head(gene_GEO,15)
    #Cross reference with the bioMart method to make sure that the conversion is correct.
```


Remove "NA" and duplicated values:
```{r}
#Removing NA values: 
gene_ARCHS4 = gene_ARCHS4_1[!is.na(gene_ARCHS4)]
gene_BRCA = gene_BRCA[!is.na(gene_BRCA)]
gene_GEO = gene_GEO[!is.na(gene_GEO)]

#Removing duplicated values:
gene_ARCHS4 = unique(gene_ARCHS4)
gene_BRCA = unique(gene_BRCA)
gene_GEO = unique(gene_GEO)
length(gene_ARCHS4)
length(gene_BRCA)
length(gene_GEO)

```

Now compare the number of overlap genes between these three data sets:
```{r}
length(intersect(gene_ARCHS4, gene_BRCA))
length(intersect(gene_ARCHS4, gene_GEO))
length((intersect(gene_BRCA, gene_GEO)))
```

```{r}
#Make Venn diagram to show overlap between 3 data sets:
venn.diagram(x = list(gene_ARCHS4, gene_BRCA, gene_GEO),
             category.names = c("ARCHS4", "BRCA", "GSE50705"),
             filename = "overlap of genes in 3 data sets.png",
             output =TRUE,
            imagetype = "png",
            scale = FALSE,
            col = "black",
            #fill = colors,
            fill = c("dodgerblue", "orchid3", "darkorange1"), 
            cat.cex = 1.2,
            margin = 0.15)


```
Next, we extract genes that are overlapped and non-overlapped between the 3 data sets to do annotations. 
```{r}
#Genes shared between all 3 data sets:
shared_3_sets = Reduce(intersect, list(gene_ARCHS4, gene_BRCA, gene_GEO))

###

#Genes shared between ARCHS4 and BRCA:
shared_ARCHS4_BRCA = intersect(gene_ARCHS4, gene_BRCA)

#Genes in ARCHS4 but not in BRCA (the elements of setdiff(x,y) are those elements in x but not in y):
differ_ARCHS4_BRCA_ARCHS4 = setdiff(x = gene_ARCHS4, y = gene_BRCA)

#Genes in BRCA but not in ARCHS4:
differ_ARCHS4_BRCA_BRCA = setdiff(x =  gene_BRCA, y = gene_ARCHS4)

###
#Genes shared between BRCA and GEO:
shared_BRCA_GEO = intersect(gene_BRCA, gene_GEO)
#Genes shared between ARCHS4 and GSE: 
shared_ARCHS4_GEO = intersect(gene_ARCHS4, gene_GEO)

#Genes in ARCHS4 but not in GEO:
differ_ARCHS4_GEO_ARCHS4 = setdiff(x = gene_ARCHS4, y = gene_GEO)

#Genes in GEO but not in ARCHS4:
differ_ARCHS4_GEO_GEO = setdiff(x =  gene_GEO, y = gene_ARCHS4)

###

#Gene in BRCA but not in GEO:
differ_BRCA_GEO_BRCA = setdiff(x = gene_BRCA, y = gene_GEO)

#Gene in GEO but not in BRCA:
differ_BRCA_GEO_GEO = setdiff(x = gene_GEO, y = gene_BRCA)

#Genes unique to GEO: 
unique_GEO = setdiff(gene_GEO, union(gene_ARCHS4, gene_BRCA))
unique_GEO[1:50]

#Genes unique to BRCA: 
unique_BRCA = setdiff(gene_BRCA, union(gene_ARCHS4, gene_GEO))

#Genes unique to ARCHS4:
unique_ARCHS4 = setdiff(gene_ARCHS4, union(gene_BRCA, gene_GEO))

#Genes overlap between ARCHS4 and GSE but not BRCA:
shared_ARCHS4_GSE_notBRCA = setdiff(x = shared_ARCHS4_GEO, y = shared_3_sets)

#Genes overlap between ARCHS4 and BRCA but not GSE:
shared_ARCHS4_BRCA_notGSE = setdiff(x = shared_ARCHS4_BRCA, y = shared_3_sets)

#Genes shared between BRCA and GSE but not ARCHS4:
shared_BRCA_GSE_notARCHS4 = setdiff(x = shared_BRCA_GEO, y = shared_3_sets)
```
Next, we export these items and do functional annotations of these lists. 
```{r}
write.csv(shared_3_sets, file = "List of genes shared between ARCHS4, GEO, and BRCA.csv")
write.csv(shared_ARCHS4_BRCA , file = "List of genes shared between ARCHS4 and BRCA.csv")
write.csv(differ_ARCHS4_BRCA_ARCHS4, file = "List of genes in ARCHS4 but not in BRCA.csv")
write.csv(shared_ARCHS4_GEO, file = "List of genes shared ARCHS4 and GEO.csv")
write.csv(unique_GEO, file = "List of genes unique to GEO.csv")
write.csv(unique_BRCA, file = "List of genes unique to BRCA.csv")
write.csv(unique_ARCHS4, file = "List of genes unique to ARCHS4.csv")
write.csv(shared_ARCHS4_GSE_notBRCA, file = "Genes shared between ARCHS4 and GSE but not BRCA.csv")
write.csv(shared_ARCHS4_BRCA_notGSE, file = "Genes shared between ARCHS4 and BRCA but not GSE.csv")
write.csv(shared_BRCA_GSE_notARCHS4, file = "Genes shared between BRCA and GSE but not ARCHS4.csv")



```

Now we install DAVID to perform annotation 
```{r}
BiocManager::install("RDAVIDWebService")
library(RDAVIDWebService)
```

```{r}
david <- DAVIDWebService$new(email= "vtran21@jhmi.edu", url="https://david.ncifcrf.gov/webservice/services/DAVIDWebService.DAVIDWebServiceHttpSoap12Endpoint/")

data(demoList1)
 result<-addList(david, demoList1, idType="AFFYMETRIX_3PRIME_IVT_ID", listName="demoList1", listType="Gene")
```

```{r}




