---
title: "Processing the data for Rank rank hypermetric test"
author: "Vy Tran"
date: "September 11, 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r library, include = FALSE}
library(BiocManager)
BiocManager::install("RRHO")
library(RRHO)
library(dplyr)

```

How many genes are overlapping between entire BRCA and GSE50705?
```{r}
MCF7 = Scaled_connectivity_for_GEO
BRCA = Scaled_connectivity_for_entire_BRCA_with_rank
colnames(MCF7) = c("Probe", "Symbol", "Scaled.connectivity.MCF7", "Rank MCF7")
colnames(BRCA) = c("Symbol", "Entrez", "Scaled.connectivity.BRCA", "Rank BRCA")

```

```{r}
MCF7 = MCF7[, c(2,3)]

#Remove genes that are duplicated and select the genes with highest scaled connectivity for each duplicated genes:
MCF7_filtered = MCF7 %>% group_by(Symbol) %>% filter(Scaled.connectivity.MCF7 == max(Scaled.connectivity.MCF7))
#Add rank of scaled connectivity:
MCF7_filtered = MCF7_filtered[order(-MCF7_filtered$Scaled.connectivity.MCF7),]
MCF7_filtered$Rank.MCF7 = rank(-MCF7_filtered$Scaled.connectivity.MCF7)

#Check and remove duplication in BRCA scaled connectivity:
BRCA = BRCA[, c(1,3)]
BRCA_filtered = BRCA %>% group_by(Symbol) %>% filter(Scaled.connectivity.BRCA == max(Scaled.connectivity.BRCA))

#Add rank of scaled connectivity:
BRCA_filtered = BRCA_filtered[order(-BRCA_filtered$Scaled.connectivity.BRCA),]
BRCA_filtered$Rank.BRCA = rank(-BRCA_filtered$Scaled.connectivity.BRCA)

rank_rank = merge(MCF7_filtered, BRCA_filtered)
```

Export the result
```{r}
write.csv(rank_rank, file = "GSE and BRCA scaled connectivity for hypermetric test_Sep_24.csv")
```
Try RRHO using the R package "RRHO". The workflow include 3 steps:
* Compute the marginal signifcance of the gene overlap for all i and j first elements of the two lists.
* Correct the marginal signifcance levels for the multiple is and js.
* Report findings using the exported signicance matrices and accom-
panying Venn diagrams.
```{r}
gene.list.MCF7 = rank_rank[, c(1,3)]
gene.list.BRCA = rank_rank[, c(1,5)]
#Compute overlap and significance:
RRHO = RRHO(gene.list.MCF7, gene.list.BRCA, BY = TRUE, alternative = "enrichment")
#Examine Nominal (-log) pvalues
lattice::levelplot(RRHO$hypermat)

# FWER corrected pvalues using 50 random permutations:
pval.testing <- pvalRRHO(RRHO, 50)
pval.testing$pval

```

```{r}
# The sampling distribution of the minimum
# of the (-log) nominal p-values:
xs<- seq(0, 10, length=100)
plot(Vectorize(pval.testing$FUN.ecdf)(xs)~xs, xlab='-log(pvalue)', ylab='ECDF', type='S')
```

```{r}
# Examine B-Y corrected pvalues.
lattice::levelplot(RRHO$hypermat.by)

```
Plot the hyper genometric overlap figures before and after correction:
```{r}
jpeg(file= "GEO and BRCA rank rank hypermetric overlap.jpeg", width = 3000, height = 3000, res = 300) 

par(mfrow=c(1,2), mar=c(5.1, 7, 4.1, 2.1))

plot.new(ttice::levelplot(RRHO$hypermat)
title(main ="a", adj=0, line=2, font=2, cex.main = 2)

lattice::levelplot(RRHO$hypermat.by)
title(main ="b", adj=0, line=2, font=2, cex.main = 2)

dev.off()
```

```
